Predictions for detecting BHNS and other double compact objects binaries with LISA

### 2021-04-29 paper/tex/sections/results.tex
update distribution grid with metallicity + more
updates to results, mainly in sc section

In this section we present our main results for the detectable LISA DCO population. We first show the distribution of the sources on the sensitivity curve before exploring the variations in the detection rate over different physics variations and analysing the parameter distributions for detectable sources in the fiducial model.

\subsection{Distribution on the sensitivity curve}

We illustrate the distribution of detectable DCOs on the sensitivity curve in Figure~\ref{fig:dcos_on_sc}. In addition to the density distributions we plot grid lines of constant distance and inspiral time to explain the shape of the distribution. The straight diagonal lines show where a binary at a fixed distance would lie on the sensitivity curve for different frequencies and the average chirp mass (annotated in each panel). We would therefore expect that if a population was entirely circular, it should be bounded approximately between the $0.1$--$30 \unit{kpc}$ lines (roughly the minimum and maximum distance to a source in the Milky Way). If we focus on the bottom panels with each individual DCO type, we see that, though this is the true for a large fraction of the population, there is a distinct subpopulation of binaries that extend downwards around $2 \times 10^{-3} \unit{Hz}$. This offshoot is composed of eccentric binaries for which the circular distance contours will not apply. We therefore also plot a line of constant distance at $30 \unit{kpc}$ for an eccentric binary with $e = 0.97$ to show the differences for eccentric sources. We also note that the peak of the density distribution coincides with the centre of the Milky Way as expected, since binaries are most likely to be formed towards the centre of the Galaxy.

We also plot vertical lines that give the inspiral time for a circular binary with the average chirp mass (annotated in each panel). From these lines, it is clear why the density distribution decreases with increasing frequency, since high frequencies correspond to short inspiral times and thus DCOs will spend less time in these regimes. We also see that the tail of the high frequency sources is more numerous near to the Galactic centre than at short distances. This is simply because there are more sources in the galactic centre and so the chances of `catching' a binary at high frequency is better.

\begin{figure*}[tp]
    \centering
    \includegraphics[width=\textwidth]{dcos_on_sc.png}
    \caption{Density distribution of detectable DCOs plotted over the LISA sensitivity curve, where the panels correspond to \textbf{top:} combined density plot for all DCO types (BHBH, BHNS and NSNS) \textbf{bottom:} three panels with individual density distribution of different DCO types. In each panel, we plot the total signal of binaries at their dominant frequency $n f_{\rm orb}$, such that $n$ is the harmonic that produces the most relative gravitational wave luminosity ($n = 2$ for circular binaries). If the density of points is below our lowest contour (2\%) then we plot the points as scatter points, where their sizes corresponds to their STROOPWAFEL weights. The inset colourbars indicate the percentage of the population represented by each contour and the annotated mass is the average chirp mass for all binaries in the panel, which is used in plotting the grid lines. We plot diagonal lines of constant distance, where the straight lines show the signal for a circular binary of average chirp mass, whilst the dashed line shows the signal for an eccentric binary with $e = 0.97$. The vertical lines indicate the inspiral time for a circular binary with the average chirp mass.}
    \label{fig:dcos_on_sc}
\end{figure*}

\subsection{Detection rates}
We find that on average, for our fiducial model, a four year LISA mission will detect \confinv{25.9}{11.1}{13.6} BHBHs, \confinv{26.7}{11.9}{14.8} BHNSs and \confinv{11.3}{6.4}{8.0} NSNSs where the error bars represent the 90\% confidence interval. Increasing the LISA mission length to ten years changes the number of detections to \confinv{42.0}{17.3}{17.3}, \confinv{44.5}{17.8}{20.7} and \confinv{19.3}{8.0}{9.7} respectively. In Figure~\ref{fig:detection_rates}, we show the expected number of LISA detections for each model variation and discuss the prominent trends in the following sections. We show the rates and uncertainties plotted in this figure in Table~\ref{tab:detection_rates}.

@@

There is a drastic decrease in detections for model \modCaseBB{} by nearly two orders of magnitude. This is because the majority of NSNS binaries are formed through case BB mass transfer and setting this mass transfer to be always unstable results in many of these binaries to merge before they could become NSNSs. As a result the total number of detections decreases, however, interestingly the remaining population represent more massive progenitors (that would not go through case BB mass transfer) and thus is skewed to higher masses and has a \textit{higher} detectable fraction (see Fig.~\ref{fig:detection_fractions}).

@@

\begin{figure*}[p]
    \centering
    \includegraphics[width=\textwidth]{dco_detections.pdf}
    \caption{The number of expected detections in the LISA mission for different DCO types and model variations. Error bars show the 1 (solid) and 2 (dotted) $\sigma$ uncertainties. The left axis and grid lines show the number of detections in a four year LISA mission and the right axis shows an approximation of the number of detections in a 10 year mission (we scale the axis by $\sqrt{T_{\rm obs}}$, see Table~\ref{tab:detection_rates} for exact rates). Each model is described in further detail in Table~\ref{tab:physics_variations} and details of the fiducial assumptions are in  Section~\ref{sec:fiducial_physics}}
    \label{fig:detection_rates}
\end{figure*}

@@

\subsubsection{Inspiral Time}
Each DCO has a strong peak at small inspiral times since higher metallicities lead to tighter binaries and thus shorter inspiral times. \todo{explain the bumps}

\begin{figure*}[htbp]
    \centering
    \includegraphics[width=\textwidth]{distribution_grid.pdf}
    \caption{Distributions for various parameters of the binaries that are detectable in a 4 year LISA mission in our fiducial model. Each line represents a kernel density estimator for the distribution and the colour denotes the double compact object type. The dotted curves in the black hole mass distribution show the primary and secondary mass distributions. In Section~\ref{sec:fiducial_distributions} we the features of the distributions.}
    \label{fig:fiducial_pdf_distributions}
\end{figure*}

>>>>

In this section we present our main results for the detectable LISA DCO population. We find that on average, for our fiducial model, a four year LISA mission will detect \confinv{34.4}{20.9}{46.6} BHBHs, \confinv{29.9}{20.1}{47.4} BHNSs and \confinv{7.7}{5.0}{10.3} NSNSs where the error bars represent the 90\% confidence interval. We first show the distribution of the sources on the sensitivity curve in Section~\ref{sec:dcos_on_sc}, before exploring the variations in the detection rate over different physics variations in Section~\ref{sec:detection_rate_analysis} and analysing the parameter distributions for detectable sources in the fiducial model in Section~\ref{sec:fiducial_distributions}.

\subsection{Distribution on the sensitivity curve}\label{sec:dcos_on_sc}

We illustrate the distribution of detectable DCOs on the sensitivity curve in Figure~\ref{fig:dcos_on_sc}. This shows that the detectable population of these massive DCOs is concentrated at comparatively lower frequencies than the LISA verification binaries (shown as stars in the top panel) and the more numerous WDWD population \citep[e.g.][]{Korol+2017}. This is expected since producing the same SNR as a BHBH, BHNS or NSNS with a relatively lower mass (circular) WDWD requires a higher frequency. This finding is in agreement with \citet{Sesana+2020} and as noted in that work, this could possibly be used to distinguish these more massive DCOs from WDWDs. There are several other notable features in these distributions and we plot grid lines of constant distance and inspiral time to help to explain the shape of the distribution.

The straight diagonal lines show where a binary at a fixed distance and with the average chirp mass (annotated in each panel) would lie on the sensitivity curve for different frequencies. We would therefore expect that if a population was entirely circular, it should be bounded approximately between the $0.1$--$30 \unit{kpc}$ lines (roughly the minimum and maximum distance to a source in the Milky Way). From inspection of the bottom panels with each individual DCO type we see that, though this is the true for a large fraction of the population, there is a distinct subpopulation of binaries that extend downwards around $2 \times 10^{-3} \unit{Hz}$. This offshoot is composed of eccentric binaries for which the circular distance contours do not apply. For instance, we plot the 90\% contour of only the circular sources in our sample over the density distribution in each of the bottom panels and it is clear that the circular subsample is bounded by the distance lines. We also plot a line of constant distance at $30 \unit{kpc}$ for an eccentric binary with $e = 0.97$ to show the differences for eccentric sources. We also note that the peak of the density distribution coincides with the centre of the Milky Way as expected, since binaries are most likely to be formed towards the centre of the Galaxy.

We also plot vertical lines that give the inspiral time for a circular binary with the average chirp mass (annotated in each panel). From these lines, it is clear why the density distribution decreases with increasing frequency, since high frequencies correspond to short inspiral times and thus DCOs will spend less time in these regimes. We also see that the tail of the high frequency sources is more numerous near to the Galactic centre than at short distances. This is simply because there are more sources in the galactic centre and so the chances of `catching' a binary at high frequency are better.

\begin{figure*}[tp]
    \centering
    \includegraphics[width=\textwidth]{dcos_on_sc.png}
    \caption{Density distribution of detectable DCOs plotted over the LISA sensitivity curve, where the panels correspond to \textbf{top:} combined density plot for all DCO types (BHBH, BHNS and NSNS) \textbf{bottom:} three panels with individual density distribution of different DCO types. In each panel, we plot the total signal of binaries at their dominant frequency $n f_{\rm orb}$, such that $n$ is the harmonic that produces the most relative gravitational wave luminosity ($n = 2$ for circular binaries). If the density of points is below our lowest contour (2\%) then we plot the points as scatter points, where their sizes corresponds to their STROOPWAFEL weights. The inset colourbars indicate the percentage of the population represented by each contour and the annotated mass is the average chirp mass for all binaries in the panel, which is used in plotting the grid lines. We plot diagonal lines of constant distance, where the straight lines show the signal for a circular binary of average chirp mass, whilst the curved line shows the signal for an eccentric binary with $e = 0.97$. The vertical lines indicate the inspiral time for a circular binary with the average chirp mass. In the top panel we overlay the LISA verification binaries from \citet{Kupfer+2018}. In the bottom panels we add the 90\% contour line for only the circular sources to show how the distribution changes without eccentric sources present.}
    \label{fig:dcos_on_sc}
\end{figure*}

\subsection{Detection rates}\label{sec:detection_rate_analysis}
We find that for our fiducial model, a four year LISA mission will detect \confinv{34.4}{20.9}{46.6} BHBHs, \confinv{29.9}{20.1}{47.4} BHNSs and \confinv{7.7}{5.0}{10.3} NSNSs, where the error bars represent the 90\% confidence interval. Increasing the LISA mission length to ten years changes the number of detections to \confinv{42.0}{17.3}{17.3}, \confinv{44.5}{17.8}{20.7} and \confinv{19.3}{8.0}{9.7} respectively. In Figure~\ref{fig:detection_rates}, we show the expected number of LISA detections for each model variation and discuss the prominent trends in the following sections. We show the rates and uncertainties plotted in this figure in Table~\ref{tab:detection_rates}.

\begin{figure*}[p]
    \centering
    \includegraphics[width=\textwidth]{dco_detections.pdf}
    \caption{The number of expected detections in the LISA mission for different DCO types and model variations. Error bars show the 50\% (solid) and 90\% (dotted) confidence intervals. The left axis and grid lines show the number of detections in a four year LISA mission and the right axis shows an approximation of the number of detections in a 10 year mission (we scale the axis by $\sqrt{T_{\rm obs}}$, see Table~\ref{tab:detection_rates} for exact rates). Each model is described in further detail in Table~\ref{tab:physics_variations} and details of the fiducial assumptions are in Section~\ref{sec:fiducial_physics}.}
    \label{fig:detection_rates}
\end{figure*}

@@

There is a drastic decrease in detections for model \modCaseBB{} by nearly two orders of magnitude. This is because the majority of NSNS binaries are formed through case BB mass transfer and setting this mass transfer to be always unstable results in many of these binaries to merge before they could become NSNSs. As a result the total number of detections decreases, however, interestingly the remaining population represent more massive progenitors (that would not go through case BB mass transfer) and thus is skewed to higher masses and has a \textit{higher} detectable fraction.

@@

@@

\subsubsection{Inspiral Time}
Each DCO has a strong peak at small inspiral times since higher metallicities lead to tighter binaries and thus shorter inspiral times. \todo{finish this}

\begin{figure*}[htbp]
    \centering
    \includegraphics[width=\textwidth]{distribution_grid_4yr.pdf}
    \caption{Distributions for various parameters of the DCOs that are detectable in a 4 year LISA mission in our fiducial model. Each panel shows the distribution of a single parameter, where the colour denotes the DCO type. We also plot the 1- and 2-$\sigma$ uncertainties (obtained via bootstrapping) with the dark and light shaded areas respectively. The first two rows (excluding metallicity) use kernel density estimators to show the distributions. The metallicity panel shows the distribution over the metallicity bins used in our population synthesis, which we show in the grid lines. The final three panels for the cumulative distribution functions for observables, normalised to the expected number of fiducial detections in a four year LISA mission. The dark shaded areas indicate regimes in which the quantity cannot be measured. The dotted lines in the angular resolution plot show the maximum angular resolution that can be covered by a single pointing of the labelled instrument. In Section~\ref{sec:fiducial_distributions} we discuss the features of the distributions.}
    \label{fig:fiducial_pdf_distributions}
\end{figure*}
